---
title: "Classifying Pitchers"
author: "Jordan Fromm"
date: "1/23/2018"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(sqldf)
library(plotly)
library(bindrcpp)
library(coin)
library(tidyr)
library(proxy)
library(vegan)
library(apcluster)
library(daff)
library(compare)
library(factoextra)
library(data.table)
library(plyr)
library(gtools)

setwd("/Users/JFromm/Desktop/Baseball/Nationals Project")
```

##Reading Data
```{r, results = 'hide'}

pitch <- read.csv("n_totalpitch.csv")
speed <- read.csv("allpitch_s.csv")
bytype <- read.csv("allpitch.csv")
```

##Creating Pitch Vector Table
```{r, results = 'hide'}
pitch_prop <- full_join(pitch, bytype, by = "pitcher", all.x = TRUE, all.y = TRUE)
colnames(pitch_prop)[4] <- "ind_pitch"
colnames(pitch_prop)[2] <- "total_pitch"
pitch_prop <- sqldf("select * from pitch_prop where ind_pitch >= 20")
pitch_prop$prop <- round(pitch_prop$ind_pitch / pitch_prop$total_pitch, 3)

pitch_prop <- pitch_prop %>% filter(pitch_type != "FO" & pitch_type != "UN" & pitch_type != "IN" & pitch_type != "PO" & pitch_type != "FA" & pitch_type != "SC")
pitch_prop <- pitch_prop[c(1,3,5)]
```

```{r, results = 'hide'}
speed <- speed %>% filter(avg_speed != "NULL") %>% filter(pitch_type != "NULL")
speed <- speed %>% filter(pitch_type != "FO" & pitch_type != "UN" & pitch_type != "IN" & pitch_type != "PO" & pitch_type != "AB")
colnames(speed)[3] <- "speed"
```

```{r, results = 'hide'}

t_CH <- pitch_prop %>% filter(pitch_type == "CH") %>% select(pitcher, prop)
s_CH <- speed %>% filter(pitch_type == "CH") %>% select(pitcher, speed)
CH <- full_join(t_CH, s_CH, by = "pitcher")
CH <- na.omit(CH)

t_CU <- pitch_prop %>% filter(pitch_type == "CU") %>% select(pitcher, prop)
s_CU <- speed %>% filter(pitch_type == "CU") %>% select(pitcher, speed)
CU <- full_join(t_CU, s_CU, by = "pitcher")
CU <- na.omit(CU)

t_EP <- pitch_prop %>% filter(pitch_type == "EP") %>% select(pitcher, prop)
s_EP <- speed %>% filter(pitch_type == "EP") %>% select(pitcher, speed)
EP <- full_join(t_EP, s_EP, by = "pitcher")
EP <- na.omit(EP)

t_FC <- pitch_prop %>% filter(pitch_type == "FC") %>% select(pitcher, prop)
s_FC <- speed %>% filter(pitch_type == "FC") %>% select(pitcher, speed)
FC <- full_join(t_FC, s_FC, by = "pitcher")
FC <- na.omit(FC)

t_FF <- pitch_prop %>% filter(pitch_type == "FF") %>% select(pitcher, prop)
s_FF <- speed %>% filter(pitch_type == "FF") %>% select(pitcher, speed)
FF <- full_join(t_FF, s_FF, by = "pitcher")
FF <- na.omit(FF)

t_FS <- pitch_prop %>% filter(pitch_type == "FS") %>% select(pitcher, prop)
s_FS <- speed %>% filter(pitch_type == "FS") %>% select(pitcher, speed)
FS <- full_join(t_FS, s_FS, by = "pitcher")
FS <- na.omit(FS)

t_FT <- pitch_prop %>% filter(pitch_type == "FT") %>% select(pitcher, prop)
s_FT <- speed %>% filter(pitch_type == "FT") %>% select(pitcher, speed)
FT <- full_join(t_FT, s_FT, by = "pitcher")
FT <- na.omit(FT)

t_KC <- pitch_prop %>% filter(pitch_type == "KC") %>% select(pitcher, prop)
s_KC <- speed %>% filter(pitch_type == "KC") %>% select(pitcher, speed)
KC <- full_join(t_KC, s_KC, by = "pitcher")
KC <- na.omit(KC)

t_KN <- pitch_prop %>% filter(pitch_type == "KN") %>% select(pitcher, prop)
s_KN <- speed %>% filter(pitch_type == "KN") %>% select(pitcher, speed)
KN <- full_join(t_KN, s_KN, by = "pitcher")
KN <- na.omit(KN)

t_SI <- pitch_prop %>% filter(pitch_type == "SI") %>% select(pitcher, prop)
s_SI <- speed %>% filter(pitch_type == "SI") %>% select(pitcher, speed)
SI <- full_join(t_SI, s_SI, by = "pitcher")
SI <- na.omit(SI)

t_SL <- pitch_prop %>% filter(pitch_type == "SL") %>% select(pitcher, prop)
s_SL <- speed %>% filter(pitch_type == "SL") %>% select(pitcher, speed)
SL <- full_join(t_SL, s_SL, by = "pitcher")
SL <- na.omit(SL)
```


```{r, warning=FALSE}
pitch_vector <- Reduce(function(x,y) merge(x,y, by = "pitcher", all = TRUE), list(CH, CU, EP, FC, FF, FS, FT, KC, KN, SI, SL))
colnames(pitch_vector) <- c("pitcher", "prop_CH", "speed_CH", "prop_CU", "speed_CU", "prop_EP", "speed_EP", "prop_FC", "speed_FC", "prop_FF", "speed_FF", "prop_FS", "speed_FS", "prop_FT", "speed_FT", "prop_KC", "speed_KC", "prop_KN", "speed_KN", "prop_SI", "speed_SI", "prop_SL", "speed_SL")
```

**Cleaning The Data**
```{r, results = 'hide'}
pitch_vector$pitcher <- as.factor(pitch_vector$pitcher)

for (i in names(pitch_vector[,c(3,5,7,9,11,13,15,17,19,21,23)])){
  pitch_vector[[i]] <- as.numeric(as.character(pitch_vector[[i]]))
}

pitch_vector[is.na(pitch_vector)] <- 0
```

##Get Pitcher Information
```{r, warning=FALSE}
batter <- read.csv("playerid.csv")
colnames(batter) <- c("batter", "lastName", "firstName", "Pos", "team")

pitchers <- read.csv("pitchers.csv")
colnames(pitchers) <- c("pitcher", "name")
pitchers$pitcher <- as.factor(pitchers$pitcher)
pitchers$name <- as.character(pitchers$name)
pitch_vector <- merge(pitch_vector, pitchers, by = "pitcher")
pitch_vector <- pitch_vector[!duplicated(pitch_vector),]

batter2 <- batter %>% filter(Pos != "P")
colnames(batter2)[1] <- "pitcher"

test3 <- merge(batter2, pitch_vector, all = TRUE)
pitch_vector <- subset(test3, is.na(test3["lastName"]))
pitch_vector <- pitch_vector[!(pitch_vector$pitcher==test3$pitcher), ]
pitch_vector <- pitch_vector[-c(2:5)]

```


##Hierarchical Clustering Tree
```{r}
pitch_vector <- pitch_vector %>% filter(name != "NULL")
test <- pitch_vector[1:30, ]
test_clust <- hclust(dist(test[, 2:23], method = "euclidean"))
plot(test_clust, labels = test$name, hang = -1)

cluster <- hclust(dist(pitch_vector[, 2:23], method = "euclidean"))
plot(cluster, labels = pitch_vector$pitcher, hang = -1)
```

##K Means Clustering
```{r, results='hide'}
set.seed(123)
fviz_nbclust(pitch_vector[, 2:23], kmeans, method = "wss")
fviz_nbclust(pitch_vector[, 2:23], kmeans, method = "silhouette")

pitchcluster <- kmeans(pitch_vector[, 2:23], 8, nstart = 20)
pitchcluster
final <- data.frame(pitchcluster$cluster, pitch_vector$pitcher)
```
##Setting Up Pitch Cluster Data

**Pitcher vs. Hitter**
```{r, results = 'hide', warning=FALSE}
pvh <- read.csv("newpvh.csv")

pvh <- pvh %>%
  mutate(uBB = ifelse(event == "Walk", 1, 0)) %>%
  mutate(IBB = ifelse(event == "Intent Walk", 1, 0)) %>%
  mutate(HBP = ifelse(event == "Hit By Pitch", 1, 0)) %>%
  mutate(SF = ifelse(event == "Sac Fly", 1, 0)) %>%
  mutate(Single = ifelse(event == "Single", 1, 0)) %>%
  mutate(Double = ifelse(event == "Double", 1, 0)) %>%
  mutate(Triple = ifelse(event == "Triple", 1, 0)) %>%
  mutate(HR = ifelse(event == "Home Run", 1, 0)) %>%
  mutate(H = ifelse((Single == 1 | Double == 1 | Triple == 1 | HR == 1), 1, 0)) %>%
  mutate(AB = ifelse((event != "Field Error" & uBB != 1 & IBB != 1 & HBP != 1 & SF != 1), 1, 0))

colnames(final)[2] <- "pitcher"
final$pitcher <- as.factor(final$pitcher)
pvh$pitcher <- as.factor(pvh$pitcher)
pvh <- full_join(final, pvh, by = "pitcher", all = TRUE)
pvh <- na.omit(pvh)
```

```{r, message=FALSE}
colnames(pvh)[1] <- "cluster"
clusters <- split(pvh, pvh$cluster)
list2env(clusters, .GlobalEnv)
```
```{r, warning=FALSE}
final <- full_join(pitch_vector, final, by = "pitcher")
final <- final[!duplicated(final$pitcher),]

final <- final[c(25, 1, 24, 2:23)]
colnames(final)[1] <- "cluster"
#final.dist <- (dist(final[, 4:25]))
#anosim(final.dist, grouping = final$cluster)

```

##Standout Batters

**Cluster 1 Overall wOBA**
```{r}
c1 <- subset(pvh, cluster == 1)
c1 <- c1[-c(2:3)]
```
```{r}
c1_PA = sqldf("select batter, count(*) as PA from c1 group by batter order by batter")
c1_Double = sqldf("select batter, count(*) as Double from c1 where Double = 1 group by batter order by batter")
c1_Triple = sqldf("select batter, count(*) as Triple from c1 where Triple = 1 group by batter order by batter")
c1_HR = sqldf("select batter, count(*) as HR from c1 where HR = 1 group by batter order by batter")
c1_uBB = sqldf("select batter, count(*) as uBB from c1 where uBB = 1 group by batter order by batter")
c1_IBB = sqldf("select batter, count(*) as IBB from c1 where IBB = 1 group by batter order by batter")
c1_HBP = sqldf("select batter, count(*) as HBP from c1 where HBP = 1 group by batter order by batter")
c1_SF = sqldf("select batter, count(*) as SF from c1 where SF = 1 group by batter order by batter")
c1_BB = sqldf("select batter, count(*) as BB from c1 where IBB = 1 or uBB = 1 group by batter order by batter")
c1_Single = sqldf("select batter, count(*) as Single from c1 where Single = 1 group by batter order by batter")
c1_AB = sqldf("select batter, count(*) as AB from c1 where AB = 1 group by batter order by batter")
c1_H = sqldf("select batter, count(*) as H from c1 where H = 1 group by batter order by batter")
```
```{r}
c1 <- merge(c1_uBB, c1_HBP, by = "batter", all.x = TRUE, all.y = TRUE)
c1 <- merge(c1, c1_Single, by = "batter", all.x = TRUE, all.y = TRUE)
c1 <- merge(c1, c1_Double, by = "batter", all.x = TRUE, all.y = TRUE)
c1 <- merge(c1, c1_Triple, by = "batter", all.x = TRUE, all.y = TRUE)
c1 <- merge(c1, c1_HR, by = "batter", all.x = TRUE, all.y = TRUE)
c1 <- merge(c1, c1_AB, by = "batter", all.x = TRUE, all.y = TRUE)
c1 <- merge(c1, c1_BB, by = "batter", all.x = TRUE, all.y = TRUE)
c1 <- merge(c1, c1_IBB, by = "batter", all.x = TRUE, all.y = TRUE)
c1 <- merge(c1, c1_SF, by = "batter", all.x = TRUE, all.y = TRUE)
c1 <- merge(c1, c1_PA, by = "batter", all.x = TRUE, all.y = TRUE)
c1[is.na(c1)] <- 0

c1$wOBA <- round(((c1$uBB*.693) + (c1$HBP*.722) + (c1$Single*.876) + (c1$Double*1.232) + (c1$Triple*1.552) + (c1$HR*1.980))
                      / (c1$AB + c1$BB - c1$IBB + c1$SF + c1$HBP), 3)
c1 <- c1 %>% filter(PA >= 62.875) %>% select(batter, PA, wOBA)
c1$cluster <- 1
c1 <- c1[c(4, 1:3)]
c1$rank <- NA
order.c1 <- order(c1$wOBA, c1$batter, decreasing = TRUE)
c1$rank[order.c1] <- 1:nrow(c1)
```

**Cluster 2 Overall wOBA**
```{r}
c2 <- subset(pvh, cluster == 2)
c2 <- c2[-c(2:3)]
```
```{r}
c2_PA = sqldf("select batter, count(*) as PA from c2 group by batter order by batter")
c2_Double = sqldf("select batter, count(*) as Double from c2 where Double = 1 group by batter order by batter")
c2_Triple = sqldf("select batter, count(*) as Triple from c2 where Triple = 1 group by batter order by batter")
c2_HR = sqldf("select batter, count(*) as HR from c2 where HR = 1 group by batter order by batter")
c2_uBB = sqldf("select batter, count(*) as uBB from c2 where uBB = 1 group by batter order by batter")
c2_IBB = sqldf("select batter, count(*) as IBB from c2 where IBB = 1 group by batter order by batter")
c2_HBP = sqldf("select batter, count(*) as HBP from c2 where HBP = 1 group by batter order by batter")
c2_SF = sqldf("select batter, count(*) as SF from c2 where SF = 1 group by batter order by batter")
c2_BB = sqldf("select batter, count(*) as BB from c2 where IBB = 1 or uBB = 1 group by batter order by batter")
c2_Single = sqldf("select batter, count(*) as Single from c2 where Single = 1 group by batter order by batter")
c2_AB = sqldf("select batter, count(*) as AB from c2 where AB = 1 group by batter order by batter")
c2_H = sqldf("select batter, count(*) as H from c2 where H = 1 group by batter order by batter")
```
```{r}
c2 <- merge(c2_uBB, c2_HBP, by = "batter", all.x = TRUE, all.y = TRUE)
c2 <- merge(c2, c2_Single, by = "batter", all.x = TRUE, all.y = TRUE)
c2 <- merge(c2, c2_Double, by = "batter", all.x = TRUE, all.y = TRUE)
c2 <- merge(c2, c2_Triple, by = "batter", all.x = TRUE, all.y = TRUE)
c2 <- merge(c2, c2_HR, by = "batter", all.x = TRUE, all.y = TRUE)
c2 <- merge(c2, c2_AB, by = "batter", all.x = TRUE, all.y = TRUE)
c2 <- merge(c2, c2_BB, by = "batter", all.x = TRUE, all.y = TRUE)
c2 <- merge(c2, c2_IBB, by = "batter", all.x = TRUE, all.y = TRUE)
c2 <- merge(c2, c2_SF, by = "batter", all.x = TRUE, all.y = TRUE)
c2 <- merge(c2, c2_PA, by = "batter", all.x = TRUE, all.y = TRUE)
c2[is.na(c2)] <- 0

c2$wOBA <- round(((c2$uBB*.693) + (c2$HBP*.722) + (c2$Single*.876) + (c2$Double*1.232) + (c2$Triple*1.552) + (c2$HR*1.980))
                 / (c2$AB + c2$BB - c2$IBB + c2$SF + c2$HBP), 3)
c2 <- c2 %>% filter(PA >= 62.875) %>% select(batter, PA, wOBA)
c2$cluster <- 2
c2 <- c2[c(4, 1:3)]
c2$rank <- NA
order.c2 <- order(c2$wOBA, c2$batter, decreasing = TRUE)
c2$rank[order.c2] <- 1:nrow(c2)
```

**Cluster 3 Overall wOBA**
```{r}
c3 <- subset(pvh, cluster == 3)
c3 <- c3[-c(2:3)]
```
```{r}
c3_PA = sqldf("select batter, count(*) as PA from c3 group by batter order by batter")
c3_Double = sqldf("select batter, count(*) as Double from c3 where Double = 1 group by batter order by batter")
c3_Triple = sqldf("select batter, count(*) as Triple from c3 where Triple = 1 group by batter order by batter")
c3_HR = sqldf("select batter, count(*) as HR from c3 where HR = 1 group by batter order by batter")
c3_uBB = sqldf("select batter, count(*) as uBB from c3 where uBB = 1 group by batter order by batter")
c3_IBB = sqldf("select batter, count(*) as IBB from c3 where IBB = 1 group by batter order by batter")
c3_HBP = sqldf("select batter, count(*) as HBP from c3 where HBP = 1 group by batter order by batter")
c3_SF = sqldf("select batter, count(*) as SF from c3 where SF = 1 group by batter order by batter")
c3_BB = sqldf("select batter, count(*) as BB from c3 where IBB = 1 or uBB = 1 group by batter order by batter")
c3_Single = sqldf("select batter, count(*) as Single from c3 where Single = 1 group by batter order by batter")
c3_AB = sqldf("select batter, count(*) as AB from c3 where AB = 1 group by batter order by batter")
c3_H = sqldf("select batter, count(*) as H from c3 where H = 1 group by batter order by batter")
```
```{r}
c3 <- merge(c3_uBB, c3_HBP, by = "batter", all.x = TRUE, all.y = TRUE)
c3 <- merge(c3, c3_Single, by = "batter", all.x = TRUE, all.y = TRUE)
c3 <- merge(c3, c3_Double, by = "batter", all.x = TRUE, all.y = TRUE)
c3 <- merge(c3, c3_Triple, by = "batter", all.x = TRUE, all.y = TRUE)
c3 <- merge(c3, c3_HR, by = "batter", all.x = TRUE, all.y = TRUE)
c3 <- merge(c3, c3_AB, by = "batter", all.x = TRUE, all.y = TRUE)
c3 <- merge(c3, c3_BB, by = "batter", all.x = TRUE, all.y = TRUE)
c3 <- merge(c3, c3_IBB, by = "batter", all.x = TRUE, all.y = TRUE)
c3 <- merge(c3, c3_SF, by = "batter", all.x = TRUE, all.y = TRUE)
c3 <- merge(c3, c3_PA, by = "batter", all.x = TRUE, all.y = TRUE)
c3[is.na(c3)] <- 0

c3$wOBA <- round(((c3$uBB*.693) + (c3$HBP*.722) + (c3$Single*.876) + (c3$Double*1.232) + (c3$Triple*1.552) + (c3$HR*1.980))
                 / (c3$AB + c3$BB - c3$IBB + c3$SF + c3$HBP), 3)
c3 <- c3 %>% filter(PA >= 62.875) %>% select(batter, PA, wOBA)
c3$cluster <- 3
c3 <- c3[c(4, 1:3)]
c3$rank <- NA
order.c3 <- order(c3$wOBA, c3$batter, decreasing = TRUE)
c3$rank[order.c3] <- 1:nrow(c3)

```

**Cluster 4 Overall wOBA**
```{r}
c4 <- subset(pvh, cluster == 4)
c4 <- c4[-c(2:3)]
```
```{r}
c4_PA = sqldf("select batter, count(*) as PA from c4 group by batter order by batter")
c4_Double = sqldf("select batter, count(*) as Double from c4 where Double = 1 group by batter order by batter")
c4_Triple = sqldf("select batter, count(*) as Triple from c4 where Triple = 1 group by batter order by batter")
c4_HR = sqldf("select batter, count(*) as HR from c4 where HR = 1 group by batter order by batter")
c4_uBB = sqldf("select batter, count(*) as uBB from c4 where uBB = 1 group by batter order by batter")
c4_IBB = sqldf("select batter, count(*) as IBB from c4 where IBB = 1 group by batter order by batter")
c4_HBP = sqldf("select batter, count(*) as HBP from c4 where HBP = 1 group by batter order by batter")
c4_SF = sqldf("select batter, count(*) as SF from c4 where SF = 1 group by batter order by batter")
c4_BB = sqldf("select batter, count(*) as BB from c4 where IBB = 1 or uBB = 1 group by batter order by batter")
c4_Single = sqldf("select batter, count(*) as Single from c4 where Single = 1 group by batter order by batter")
c4_AB = sqldf("select batter, count(*) as AB from c4 where AB = 1 group by batter order by batter")
c4_H = sqldf("select batter, count(*) as H from c4 where H = 1 group by batter order by batter")
```
```{r}
c4 <- merge(c4_uBB, c4_HBP, by = "batter", all.x = TRUE, all.y = TRUE)
c4 <- merge(c4, c4_Single, by = "batter", all.x = TRUE, all.y = TRUE)
c4 <- merge(c4, c4_Double, by = "batter", all.x = TRUE, all.y = TRUE)
c4 <- merge(c4, c4_Triple, by = "batter", all.x = TRUE, all.y = TRUE)
c4 <- merge(c4, c4_HR, by = "batter", all.x = TRUE, all.y = TRUE)
c4 <- merge(c4, c4_AB, by = "batter", all.x = TRUE, all.y = TRUE)
c4 <- merge(c4, c4_BB, by = "batter", all.x = TRUE, all.y = TRUE)
c4 <- merge(c4, c4_IBB, by = "batter", all.x = TRUE, all.y = TRUE)
c4 <- merge(c4, c4_SF, by = "batter", all.x = TRUE, all.y = TRUE)
c4 <- merge(c4, c4_PA, by = "batter", all.x = TRUE, all.y = TRUE)
c4[is.na(c4)] <- 0

c4$wOBA <- round(((c4$uBB*.693) + (c4$HBP*.722) + (c4$Single*.876) + (c4$Double*1.232) + (c4$Triple*1.552) + (c4$HR*1.980))
                 / (c4$AB + c4$BB - c4$IBB + c4$SF + c4$HBP), 3)
c4 <- c4 %>% filter(PA >= 62.875) %>% select(batter, PA, wOBA)
c4$cluster <- 4
c4 <- c4[c(4, 1:3)]
c4$rank <- NA
order.c4 <- order(c4$wOBA, c4$batter, decreasing = TRUE)
c4$rank[order.c4] <- 1:nrow(c4)
```

**Cluster 5 Overall wOBA**
```{r}
c5 <- subset(pvh, cluster == 5)
c5 <- c5[-c(2:3)]
```
```{r}
c5_PA = sqldf("select batter, count(*) as PA from c5 group by batter order by batter")
c5_Double = sqldf("select batter, count(*) as Double from c5 where Double = 1 group by batter order by batter")
c5_Triple = sqldf("select batter, count(*) as Triple from c5 where Triple = 1 group by batter order by batter")
c5_HR = sqldf("select batter, count(*) as HR from c5 where HR = 1 group by batter order by batter")
c5_uBB = sqldf("select batter, count(*) as uBB from c5 where uBB = 1 group by batter order by batter")
c5_IBB = sqldf("select batter, count(*) as IBB from c5 where IBB = 1 group by batter order by batter")
c5_HBP = sqldf("select batter, count(*) as HBP from c5 where HBP = 1 group by batter order by batter")
c5_SF = sqldf("select batter, count(*) as SF from c5 where SF = 1 group by batter order by batter")
c5_BB = sqldf("select batter, count(*) as BB from c5 where IBB = 1 or uBB = 1 group by batter order by batter")
c5_Single = sqldf("select batter, count(*) as Single from c5 where Single = 1 group by batter order by batter")
c5_AB = sqldf("select batter, count(*) as AB from c5 where AB = 1 group by batter order by batter")
c5_H = sqldf("select batter, count(*) as H from c5 where H = 1 group by batter order by batter")
```
```{r}
c5 <- merge(c5_uBB, c5_HBP, by = "batter", all.x = TRUE, all.y = TRUE)
c5 <- merge(c5, c5_Single, by = "batter", all.x = TRUE, all.y = TRUE)
c5 <- merge(c5, c5_Double, by = "batter", all.x = TRUE, all.y = TRUE)
c5 <- merge(c5, c5_Triple, by = "batter", all.x = TRUE, all.y = TRUE)
c5 <- merge(c5, c5_HR, by = "batter", all.x = TRUE, all.y = TRUE)
c5 <- merge(c5, c5_AB, by = "batter", all.x = TRUE, all.y = TRUE)
c5 <- merge(c5, c5_BB, by = "batter", all.x = TRUE, all.y = TRUE)
c5 <- merge(c5, c5_IBB, by = "batter", all.x = TRUE, all.y = TRUE)
c5 <- merge(c5, c5_SF, by = "batter", all.x = TRUE, all.y = TRUE)
c5 <- merge(c5, c5_PA, by = "batter", all.x = TRUE, all.y = TRUE)
c5[is.na(c5)] <- 0

c5$wOBA <- round(((c5$uBB*.693) + (c5$HBP*.722) + (c5$Single*.876) + (c5$Double*1.232) + (c5$Triple*1.552) + (c5$HR*1.980))
                 / (c5$AB + c5$BB - c5$IBB + c5$SF + c5$HBP), 3)
c5 <- c5 %>% filter(PA >= 62.875) %>% select(batter, PA, wOBA)
c5$cluster <- 5
c5 <- c5[c(4, 1:3)]
c5$rank <- NA
order.c5 <- order(c5$wOBA, c5$batter, decreasing = TRUE)
c5$rank[order.c5] <- 1:nrow(c5)
```

**Cluster 6 Overall wOBA**
```{r}
c6 <- subset(pvh, cluster == 6)
c6 <- c6[-c(2:3)]
```
```{r}
c6_PA = sqldf("select batter, count(*) as PA from c6 group by batter order by batter")
c6_Double = sqldf("select batter, count(*) as Double from c6 where Double = 1 group by batter order by batter")
c6_Triple = sqldf("select batter, count(*) as Triple from c6 where Triple = 1 group by batter order by batter")
c6_HR = sqldf("select batter, count(*) as HR from c6 where HR = 1 group by batter order by batter")
c6_uBB = sqldf("select batter, count(*) as uBB from c6 where uBB = 1 group by batter order by batter")
c6_IBB = sqldf("select batter, count(*) as IBB from c6 where IBB = 1 group by batter order by batter")
c6_HBP = sqldf("select batter, count(*) as HBP from c6 where HBP = 1 group by batter order by batter")
c6_SF = sqldf("select batter, count(*) as SF from c6 where SF = 1 group by batter order by batter")
c6_BB = sqldf("select batter, count(*) as BB from c6 where IBB = 1 or uBB = 1 group by batter order by batter")
c6_Single = sqldf("select batter, count(*) as Single from c6 where Single = 1 group by batter order by batter")
c6_AB = sqldf("select batter, count(*) as AB from c6 where AB = 1 group by batter order by batter")
c6_H = sqldf("select batter, count(*) as H from c6 where H = 1 group by batter order by batter")
```
```{r}
c6 <- merge(c6_uBB, c6_HBP, by = "batter", all.x = TRUE, all.y = TRUE)
c6 <- merge(c6, c6_Single, by = "batter", all.x = TRUE, all.y = TRUE)
c6 <- merge(c6, c6_Double, by = "batter", all.x = TRUE, all.y = TRUE)
c6 <- merge(c6, c6_Triple, by = "batter", all.x = TRUE, all.y = TRUE)
c6 <- merge(c6, c6_HR, by = "batter", all.x = TRUE, all.y = TRUE)
c6 <- merge(c6, c6_AB, by = "batter", all.x = TRUE, all.y = TRUE)
c6 <- merge(c6, c6_BB, by = "batter", all.x = TRUE, all.y = TRUE)
c6 <- merge(c6, c6_IBB, by = "batter", all.x = TRUE, all.y = TRUE)
c6 <- merge(c6, c6_SF, by = "batter", all.x = TRUE, all.y = TRUE)
c6 <- merge(c6, c6_PA, by = "batter", all.x = TRUE, all.y = TRUE)
c6[is.na(c6)] <- 0

c6$wOBA <- round(((c6$uBB*.693) + (c6$HBP*.722) + (c6$Single*.876) + (c6$Double*1.232) + (c6$Triple*1.552) + (c6$HR*1.980))
                 / (c6$AB + c6$BB - c6$IBB + c6$SF + c6$HBP), 3)
c6 <- c6 %>% filter(PA >= 62.875) %>% select(batter, PA, wOBA)
c6$cluster <- 6
c6 <- c6[c(4, 1:3)]
c6$rank <- NA
order.c6 <- order(c6$wOBA, c6$batter, decreasing = TRUE)
c6$rank[order.c6] <- 1:nrow(c6)
```

**Cluster 7 Overall wOBA**
```{r}
c7 <- subset(pvh, cluster == 7)
c7 <- c7[-c(2:3)]
```
```{r}
c7_PA = sqldf("select batter, count(*) as PA from c7 group by batter order by batter")
c7_Double = sqldf("select batter, count(*) as Double from c7 where Double = 1 group by batter order by batter")
c7_Triple = sqldf("select batter, count(*) as Triple from c7 where Triple = 1 group by batter order by batter")
c7_HR = sqldf("select batter, count(*) as HR from c7 where HR = 1 group by batter order by batter")
c7_uBB = sqldf("select batter, count(*) as uBB from c7 where uBB = 1 group by batter order by batter")
c7_IBB = sqldf("select batter, count(*) as IBB from c7 where IBB = 1 group by batter order by batter")
c7_HBP = sqldf("select batter, count(*) as HBP from c7 where HBP = 1 group by batter order by batter")
c7_SF = sqldf("select batter, count(*) as SF from c7 where SF = 1 group by batter order by batter")
c7_BB = sqldf("select batter, count(*) as BB from c7 where IBB = 1 or uBB = 1 group by batter order by batter")
c7_Single = sqldf("select batter, count(*) as Single from c7 where Single = 1 group by batter order by batter")
c7_AB = sqldf("select batter, count(*) as AB from c7 where AB = 1 group by batter order by batter")
c7_H = sqldf("select batter, count(*) as H from c7 where H = 1 group by batter order by batter")
```
```{r}
c7 <- merge(c7_uBB, c7_HBP, by = "batter", all.x = TRUE, all.y = TRUE)
c7 <- merge(c7, c7_Single, by = "batter", all.x = TRUE, all.y = TRUE)
c7 <- merge(c7, c7_Double, by = "batter", all.x = TRUE, all.y = TRUE)
c7 <- merge(c7, c7_Triple, by = "batter", all.x = TRUE, all.y = TRUE)
c7 <- merge(c7, c7_HR, by = "batter", all.x = TRUE, all.y = TRUE)
c7 <- merge(c7, c7_AB, by = "batter", all.x = TRUE, all.y = TRUE)
c7 <- merge(c7, c7_BB, by = "batter", all.x = TRUE, all.y = TRUE)
c7 <- merge(c7, c7_IBB, by = "batter", all.x = TRUE, all.y = TRUE)
c7 <- merge(c7, c7_SF, by = "batter", all.x = TRUE, all.y = TRUE)
c7 <- merge(c7, c7_PA, by = "batter", all.x = TRUE, all.y = TRUE)
c7[is.na(c7)] <- 0

c7$wOBA <- round(((c7$uBB*.693) + (c7$HBP*.722) + (c7$Single*.876) + (c7$Double*1.232) + (c7$Triple*1.552) + (c7$HR*1.980))
                 / (c7$AB + c7$BB - c7$IBB + c7$SF + c7$HBP), 3)
c7 <- c7 %>% filter(PA >= 62.875) %>% select(batter, PA, wOBA)
c7$cluster <- 7
c7 <- c7[c(4, 1:3)]
c7$rank <- NA
order.c7 <- order(c7$wOBA, c7$batter, decreasing = TRUE)
c7$rank[order.c7] <- 1:nrow(c7)
```

**Cluster 8 Overall wOBA**
```{r}
c8 <- subset(pvh, cluster == 8)
c8 <- c8[-c(2:3)]
```
```{r}
c8_PA = sqldf("select batter, count(*) as PA from c8 group by batter order by batter")
c8_Double = sqldf("select batter, count(*) as Double from c8 where Double = 1 group by batter order by batter")
c8_Triple = sqldf("select batter, count(*) as Triple from c8 where Triple = 1 group by batter order by batter")
c8_HR = sqldf("select batter, count(*) as HR from c8 where HR = 1 group by batter order by batter")
c8_uBB = sqldf("select batter, count(*) as uBB from c8 where uBB = 1 group by batter order by batter")
c8_IBB = sqldf("select batter, count(*) as IBB from c8 where IBB = 1 group by batter order by batter")
c8_HBP = sqldf("select batter, count(*) as HBP from c8 where HBP = 1 group by batter order by batter")
c8_SF = sqldf("select batter, count(*) as SF from c8 where SF = 1 group by batter order by batter")
c8_BB = sqldf("select batter, count(*) as BB from c8 where IBB = 1 or uBB = 1 group by batter order by batter")
c8_Single = sqldf("select batter, count(*) as Single from c8 where Single = 1 group by batter order by batter")
c8_AB = sqldf("select batter, count(*) as AB from c8 where AB = 1 group by batter order by batter")
c8_H = sqldf("select batter, count(*) as H from c8 where H = 1 group by batter order by batter")
```
```{r}
c8 <- merge(c8_uBB, c8_HBP, by = "batter", all.x = TRUE, all.y = TRUE)
c8 <- merge(c8, c8_Single, by = "batter", all.x = TRUE, all.y = TRUE)
c8 <- merge(c8, c8_Double, by = "batter", all.x = TRUE, all.y = TRUE)
c8 <- merge(c8, c8_Triple, by = "batter", all.x = TRUE, all.y = TRUE)
c8 <- merge(c8, c8_HR, by = "batter", all.x = TRUE, all.y = TRUE)
c8 <- merge(c8, c8_AB, by = "batter", all.x = TRUE, all.y = TRUE)
c8 <- merge(c8, c8_BB, by = "batter", all.x = TRUE, all.y = TRUE)
c8 <- merge(c8, c8_IBB, by = "batter", all.x = TRUE, all.y = TRUE)
c8 <- merge(c8, c8_SF, by = "batter", all.x = TRUE, all.y = TRUE)
c8 <- merge(c8, c8_PA, by = "batter", all.x = TRUE, all.y = TRUE)
c8[is.na(c8)] <- 0

c8$wOBA <- round(((c8$uBB*.693) + (c8$HBP*.722) + (c8$Single*.876) + (c8$Double*1.232) + (c8$Triple*1.552) + (c8$HR*1.980))
                 / (c8$AB + c8$BB - c8$IBB + c8$SF + c8$HBP), 3)
c8 <- c8 %>% filter(PA >= 62.875) %>% select(batter, PA, wOBA)
c8$cluster <- 8
c8 <- c8[c(4, 1:3)]
c8$rank <- NA
order.c8 <- order(c8$wOBA, c8$batter, decreasing = TRUE)
c8$rank[order.c8] <- 1:nrow(c8)
```

**Merging Data Frames**

```{r}
overall <- Reduce(function(x,y) merge(x,y, all = TRUE), list(c1, c2, c3, c4, c5, c6, c7, c8))
overall <- na.omit(overall)
overall <- sqldf("select cluster, batter, PA, wOBA from overall order by batter, cluster")
```
```{r}
players <- inner_join(c1, c2, by = "batter")
players <- inner_join(players, c3, by = "batter")
players <- inner_join(players, c4, by = "batter")
players <- inner_join(players, c5, by = "batter")
players <- inner_join(players, c6, by = "batter")
players <- inner_join(players, c7, by = "batter")
players <- inner_join(players, c8, by = "batter")
```
```{r}
overall <- overall[overall$batter %in% players$batter,]
```
```{r}
maxwOBA <- sqldf("select batter, max(wOBA) as max_wOBA from overall group by batter")
minwOBA <- sqldf("select batter, min(wOBA) as min_wOBA from overall group by batter")
maxmin <- merge(maxwOBA, minwOBA)
maxmin$range <- maxmin$max_wOBA-maxmin$min_wOBA
maxmin <- maxmin[c(1, 4)]

meanwOBA <- sqldf("select batter, avg(wOBA) as mean_wOBA from overall group by batter")
meanmin <- merge(meanwOBA, minwOBA)
meanmin$diff <- round(meanmin$mean_wOBA-meanmin$min_wOBA, 3)
meanmin <- meanmin[c(1,4)]

plot_data <- merge(maxmin, meanmin)
plot_data$total <- plot_data$range + plot_data$diff
plot_data$rank <- NA
plot.order <- order(plot_data$total, plot_data$batter, decreasing = TRUE)
plot_data$rank[plot.order] <- 1:nrow(plot_data)

plot_data2 <- merge(maxmin, meanwOBA)
plot_data2$mean_wOBA <- round(plot_data2$mean_wOBA, 3)
colnames(plot_data2)[3] <- "mean"
plot_data2$total <- plot_data2$range + plot_data2$mean
plot_data2$rank <- NA
plot2.order <- order(plot_data2$total, plot_data2$batter, decreasing = TRUE)
plot_data2$rank[plot2.order] <- 1:nrow(plot_data2)
```
```{r}
newplot <- ggplot(plot_data, aes(x = range, y = diff)) +
  geom_point() +
  labs(x = "Range", y = "Mean - Min")
ggplotly(newplot)
```
```{r, message=FALSE, warning=FALSE}
newplot2 <- ggplot(plot_data2, aes(x = range, y = mean)) +
  geom_point() +
  labs(x = "Range", y = "Mean") +
  geom_segment(aes(x = 0, xend = quantile(plot_data2$range, 0.1), y = quantile(plot_data2$mean, 0.1), yend = quantile(plot_data2$mean, 0.1))) +
  geom_segment(aes(x = quantile(plot_data2$range, 0.1), xend = quantile(plot_data2$range, 0.1), y = 0.25, yend = quantile(plot_data2$mean, 0.1))) +
  geom_segment(x = 0, xend = quantile(plot_data2$range, 0.1), y = quantile(plot_data2$mean, 0.9), yend = quantile(plot_data2$mean, 0.9)) +
  geom_segment(aes(x = quantile(plot_data2$range, 0.1), xend = quantile(plot_data2$range, 0.1), y = quantile(plot_data2$mean, 0.9), yend = 0.45)) +
  geom_segment(aes(x = quantile(plot_data2$range, 0.9), xend = 0.25, y = quantile(plot_data2$mean, 0.9), yend = quantile(plot_data2$mean, 0.9))) + 
  geom_segment(aes(x = quantile(plot_data2$range, 0.9), xend = quantile(plot_data2$range, 0.9), y = quantile(plot_data2$mean, 0.9), yend = 0.45)) +
  geom_segment(aes(x = quantile(plot_data2$range, 0.9), xend = 0.25, y = quantile(plot_data2$mean, 0.1), yend = quantile(plot_data2$mean, 0.1))) +
  geom_segment(aes(x = quantile(plot_data2$range, 0.9), xend = quantile(plot_data2$range, 0.9), y = 0.25, yend = quantile(plot_data2$mean, 0.1))) +
  ggtitle("Batter Mean wOBA vs. Range of wOBA \n (Amongst All Clusters)") +
  theme(plot.title = element_text(hjust = 0.5))
ggplotly(newplot2)
```

**Correlation**
```{r, comment = NA}
correlation <- Reduce(function(x,y) merge(x,y, all = TRUE), list(c1, c2, c3, c4, c5, c6, c7, c8))
correlation <- correlation[correlation$batter %in% overall$batter,]
correlation <- as.data.table(correlation)
correlation$cluster <- as.character(correlation$cluster)
correlation <- correlation[,c(1,5)]

setkey(correlation, "cluster")
clust = unique(correlation$cluster)
nclust = length(clust)
pairs = combinations(nclust,2,clust)

p1 <- correlation[pairs[,1],,allow.cartesian = TRUE]
p2 <- correlation[pairs[,2],,allow.cartesian = TRUE]
correlation <- p1[, c("clust2", "rank2"):= p2]
correlation <- p1[,cor(rank, rank2), by = list(cluster, clust2)]
colnames(correlation) <- c("clusterA", "clusterB", "cor")
correlation
```

**Interesting Players**
```{r}
bathand <- read.csv("batterwithhand.csv")
colnames(bathand) <- c("batter", "name", "stand")
bathand <- merge(batter, bathand, by = "batter")
bathand <- bathand[,-c(5:6)]

topleft <- sqldf("select * from plot_data2 where mean >= 0.357 and range <= 0.066")
topright <- sqldf("select * from plot_data2 where range >= 0.164 and mean >= 0.357")
bottomright <- sqldf("select * from plot_data2 where mean <= 0.291 and range >= 0.164")

topleft <- merge(topleft, bathand, by = "batter")
topright <- merge(topright, bathand, by = "batter")
bottomright <- merge(bottomright, bathand, by = "batter")
```
